---
title: "SNOTEL Analysis"
author: "Kyle Bocinsky"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is just a brief example of one of the `mcor` package's capabilities---namely, accessing and processing the SNOTEL dataset from the Natural Resources Conservation Service of the USDA. In this vignette, we'll create a map-based dashboard for the SNOTEL sites. We will:

- Download the locations and metadata of all SNOTEL sites in Montana
- Download the normal data record for each station
- Download the most recent season's data for each station
- Map the stations on a leaflet map
- Graph the SWE accumulation in a Plotly popup

## Accessing the SNOTEL data
```{r snotel-data}
library(mcor)
library(magrittr)
library(tidyverse)
library(sf)
library(leaflet)

snotel_inventory <- mco_get_snotel_inventory() %>%
  dplyr::filter(`State Code` == "MT") %>%
  dplyr::filter(`Start Date` <= as.Date("1981-01-01")) %>%
  dplyr::mutate(station = paste0(`Station Id`,":",
                                 `State Code`,":",
                                 `Network Code`))

# A map of snotel sites in Montana
leaflet::leaflet(width = "100%", 
                 height = "100%") %>%
  leaflet::addProviderTiles("OpenTopoMap", group = "Topo") %>%
  leaflet::addProviderTiles("OpenStreetMap.BlackAndWhite", group = "OpenStreetMap") %>%
  leaflet::addProviderTiles("Esri.WorldImagery", group = "Satellite") %>%
  leaflet::addProviderTiles("Stamen.TonerLines",
                            group = "Satellite"
  ) %>%
  leaflet::addProviderTiles("Stamen.TonerLabels",
                            group = "Satellite"
  ) %>%
  leaflet::addCircleMarkers(data = snotel_inventory,
                            label = ~htmltools::htmlEscape(`Station Name`),
                            radius = 5,
                            stroke = FALSE,
                            fillOpacity = 0.75,
                            col = "white") %>%
  leaflet::addLayersControl(
    baseGroups = c("Topo",
                   "OpenStreetMap",
                   "Satellite"),
    options = leaflet::layersControlOptions(collapsed = FALSE),
    position = "topleft"
  )


snotel_normals <- readr::read_csv(paste0("https://wcc.sc.egov.usda.gov/reportGenerator/view_csv/customMultipleStationReport/daily/start_of_period/",
                                         paste0(snotel_inventory$station %>% unique,
                                                collapse = "%7C"),
                                         # "%7C","state=%22MT%22",
                                         # "%20AND%20",
                                         "network=%22SNTL%22",
                                         # "%20AND%20element=%22WTEQ%22",
                                         # "%20AND%20outServiceDate=%222100-01-01%22",
                                         "%7Cname/","1981-01-01",",","2010-12-31","/stationId,",
                                         "WTEQ::value,",
                                         "?fitToScreen=false"),
                                  comment = "#",
                                  col_types = readr::cols(
                                    `Date` = readr::col_date(format = ""),
                                    `Station Id` = readr::col_integer(),
                                    `Snow Water Equivalent (in) Start of Day Values` = readr::col_double()
                                  )) %>%
  dplyr::mutate(`Station Id` = factor(`Station Id`)) %>%
  dplyr::rename(SWE = `Snow Water Equivalent (in) Start of Day Values`)

snotel_normals_quantiles <- snotel_normals %>%
  dplyr::mutate(doy = lubridate::yday(Date) %>% 
                  as.integer()) %>%
  dplyr::group_by(doy,`Station Id`) %>%
  dplyr::arrange(doy,`Station Id`) %>%
  dplyr::summarise(mean = mean(SWE, na.rm = TRUE),
                   `0%` = quantile(SWE, probs = 0, na.rm = TRUE),
                   `10%` = quantile(SWE, probs = 0.1, na.rm = TRUE),
                   `30%` = quantile(SWE, probs = 0.3, na.rm = TRUE),
                   `50%` = quantile(SWE, probs = 0.5, na.rm = TRUE),
                   `70%` = quantile(SWE, probs = 0.7, na.rm = TRUE),
                   `90%` = quantile(SWE, probs = 0.9, na.rm = TRUE),
                   `100%` = quantile(SWE, probs = 1, na.rm = TRUE)) %>%
  dplyr::filter(doy != 366)

snotel_2018 <- readr::read_csv(paste0("https://wcc.sc.egov.usda.gov/reportGenerator/view_csv/customMultipleStationReport/daily/start_of_period/",
                                      paste0(snotel_inventory$station %>% unique,
                                             collapse = "%7C"),
                                      "%7C","state=%22MT%22",
                                      "%20AND%20",
                                      "network=%22SNTL%22",
                                      "%20AND%20element=%22WTEQ%22",
                                      "%20AND%20outServiceDate=%222100-01-01%22",
                                      "%7Cname/","2017-10-01",",",Sys.Date(),"/stationId,",
                                      "WTEQ::value,",
                                      "?fitToScreen=false"),
                               comment = "#",
                               col_types = readr::cols(
                                 `Date` = readr::col_date(format = ""),
                                 `Station Id` = readr::col_integer(),
                                 `Snow Water Equivalent (in) Start of Day Values` = readr::col_double()
                               )) %>%
  dplyr::mutate(`Station Id` = factor(`Station Id`),
                doy = lubridate::yday(Date) %>% 
                  as.integer()) %>%
  dplyr::rename(SWE = `Snow Water Equivalent (in) Start of Day Values`)


snotel_2018_test <- snotel_2018 %>%
  dplyr::filter(`Station Id` == 328)

snotel_normals_quantiles_test <- snotel_normals_quantiles %>%
  dplyr::filter(`Station Id` == 328)

# dowy <- length(snotel_2018_test$SWE %>% na.omit())

snotel_2018_test %<>%
  dplyr::right_join(snotel_normals_quantiles_test,
                   by = c("doy",
                          "Station Id")) %>%
  dplyr::mutate(dowy = doy %>%
                  magrittr::subtract(273) %>%
                  ifelse(. <= 0, .+365, .) %>%
                  as.integer(),
                Date = as.Date(dowy-1, origin = "2017-10-01")) %>%
  dplyr::arrange(Date)

(snotel_2018_test %>%
  dplyr::ungroup() %>%
  ggplot(aes(x = Date)) +
  geom_ribbon(aes(ymin = `0%`,
                  ymax = `100%`),
              fill = "grey80") +
  geom_ribbon(aes(ymin = `10%`,
                  ymax = `90%`),
              fill = "grey65") + 
  geom_ribbon(aes(ymin = `30%`,
                  ymax = `70%`),
              fill = "grey50") +
  geom_line(aes(y = `50%`),
            color = "black") +
    geom_line(aes(y = `mean`),
            color = "black",
            linetype = 3) +
  geom_line(aes(y = SWE),
            color = "red",
            size = 1) +
      # geom_line(aes(y = `0% NEP`),
      #       color = "dodgerblue",
      #       size = 1) +
    ylab("Snow Water Equivalent (in)") +
    scale_x_date(labels = function(x) format(x, "%d-%b"))) %>%
  plotly::ggplotly()



```
